---
title: "Consumer Behaviour Analysis"
subtitle: "Proposal"
author: 
  - name: "Pattern Pioneers - Vishal, Joel, Pranshu, Shashwat, Bharath"
    affiliations:
      - name: "School of Information, University of Arizona"
description: ""
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
categories: 
  - Data Mining
execute:
  warning: false
---

## Load required packages

```{r load_pkgs, message=FALSE, output=FALSE}
#| code-fold: true
#| code-summary: "Installed Packages"

### GETTING THE LIBRARIES
if (!require(pacman))
  install.packages(pacman)

pacman::p_load(tidyverse,   # Data wrangling
               dlookr,      # Exploratory Data Analysis
               formattable, # Present neat table format
               gt,          # Alternating formatting for the tables
               gtsummary)   # Summary for the tables
```

## Goal

Our main motivation for selecting the dataset **Amazon consumer Behaviour Dataset** that we came across in Kaggle was to unveil some customer insights which can be used for enhancing the customer experience or the improving the business implementation after analyzing this dataset.

## Dataset

```{r}
#| label: load-dataset
#| message: false
#| code-fold: true

# Using the original data
# Loading the csv into a variable using read_csv

data <- read_csv("data/Amazon_Customer_Behavior_Survey.csv")

# Removing unwanted column like "timestamp".

data <- data %>% select(-Timestamp)
```

This is a dataset collected from [kaggle](https://www.kaggle.com/datasets/swathiunnikrishnan/amazon-consumer-behaviour-dataset/data) for analyzing the behavioral analysis of Amazon's consumers consists of a comprehensive collection of customer interactions, browsing patterns within the Amazon ecosystem. It includes a wide range of variables such as customer demographics, user interaction, and reviews. The dataset aims to provide insights into customer preferences, shopping habits, and decision-making processes on the Amazon platform. By analyzing this dataset, researchers and analysts can gain a deeper understanding of consumer behavior, identify trends, optimize marketing strategies, and improve the overall customer experience on Amazon. The Dataset contains `N=602` observations.

## Examine data

Using dlookr's `describe()` and `diagnose()` - some basic EDA

```{r}
#| label: data-summary-describe
#| message: false
#| code-fold: true
# Summary statistics of numerical column

data |> 
    describe() |>
  formattable()
```

```{r}
#| label: data-summary-diagnose 
#| message: false
#| code-fold: true

# using Diagnose from dlookr to look for column summary
data |> 
  diagnose() |>
  formattable()
```

Checking the number of rows and columns with `nrow` and `ncol`:

```{r}
#| label: No-of-row
#| message: false
#| code-fold: true

# Number of rows in the data
nrow(data)
```

-   So we have totally 602 data points in the Amazon Consumer behavior dataset. One important point to note here is that some rows contain multiple entries for

```{r}
#| label: No-of-col 
#| message: false
#| code-fold: true
# Number of columns in the data
ncol(data)
```

-   And we have 22 columns in the dataset.

## Categorical variable summary

Using `gtsummary` for table summary (`tbl_summary()`)of selected categorical columns:

```{r}
#| label: gt-summary
#| message: false
#| code-fold: true
# Selecting the required columns for summary 
 new_data <-data %>% select(Browsing_Frequency,Purchase_Frequency,Purchase_Categories)
  
# Using gtsummary
 
new_data |>
  gtsummary::tbl_summary()
```

## Questions

### Question 1

In our first question "**What are the factors influencing the customer's decision to abandon a purchase in their cart on Amazon?**" we are attempting to understand the reasons behind the customer abandoning the purchase in their cart for increasing the **conversion rate(**the percentage of users who actually complete a purchase**)** for amazon. It will also help us in enhancing the customer experience by making the application or the website more user-friendly and intuitive so that the user is able to find the right product and proceed to complete his purchase in an effortless manner.

### Question 2

For our second question "**Which demographic (on the basis of gender and age) is most likely to purchase a particular product category?**" we attempt to determine the demographic which is most likely to purchase a particular product category on the basis of their age and gender which will help companies to tailor their marketing strategies so that their messages are able to reach the right group of customers, leading to cost optimization of their marketing budget. By identifying the right demographic to target amazon can gain a competitive advantage by attracting a larger share of the target audience. This will also lead to higher levels of customer satisfaction which will result in customer loyalty and recurring purchases.

## Analysis plan

### Approach for question 1

Our analysis will commence with a precise exploratory data analysis (EDA) using R, where we'll focus on key variables such as `Purchase_Frequency`, `Product_Search_Method`, and `Customer_Reviews_Importance`, to unearth their potential impact on cart abandonment. During the data cleaning stage, we will ensure that columns like `age`, `Gender`, and `Browsing_Frequency` are correctly formatted and free of missing values. Subsequently, statistical techniques---such as logistic regression or decision trees---will be applied to assess the influence of factors like `Personalized_Recommendation_Frequency` and `Search_Result_Exploration` on cart abandonment behavior.

For visualization, we will harness the capabilities of R's ggplot2 package to create comprehensive and interpretable graphics, such as correlation heat-maps and bar charts, showcasing the interplay between cart abandonment and various customer behavior metrics. This meticulous approach is designed to provide us with robust insights, empowering us to enhance the shopping experience and curtail cart abandonment rates effectively.

### Approach for question 2

We will further continue our analysis where we will be focusing majorly on the key features such as `age`, `Gender`, `Purchase_Categories` and `Cart_Completion_Frequency` to predict which product is of particular interest to a demography. Again, we will make sure the data is clean for the concerned features and address any missing or incorrect data.

Post that we will identify important summary statistics for key variables such as `Purchase_Categories`. Once the data is clean, we will employ Statistical Analysis tools to assess the likelihood of purchasing product categories based on demographic variables (age and gender). For visualization, we will be using heat-maps and bar charts to illustrate the relationship between the target variables.

### Variables of focus for both questions

| Variable                              | Description                                                                                |
|----------------------|--------------------------------------------------|
| Purchase_Frequency                    | How frequently does the user make purchases on Amazon?                                     |
| Product_Search_Method                 | How does the user search for products on Amazon?                                           |
| Customer_Reviews_Importance           | How important are customer reviews in users decision-making process?                       |
| age                                   | age                                                                                        |
| Gender                                | gender                                                                                     |
| Browsing_Frequency                    | How often does the user browse Amazon's website or app?                                    |
| Purchase_Categories                   | What product categories does the user typically purchase on Amazon?                        |
| Cart_Completion_Frequency             | How often do user complete the purchase after adding products to their cart?               |
| Personalized_Recommendation_Frequency | How often do user receive personalized product recommendations from Amazon?                |
| Search_Result_Exploration             | Does the user tend to explore multiple pages of search results or focus on the first page? |

## Organization

### Plan of Attack

|               Week                | Weekly Tasks                                                           | Persons in Charge |  Backup  |
|:------------:|--------------------------------|:-----------:|:-----------:|
|       until November 8^th^        | Explore and finalize the dataset and the problem statements            |     Everyone      | Everyone |
|                \-                 | Complete the proposal and assign some high-level tasks                 |     Everyone      | Everyone |
|     November 9^th^ to 15^th^      | Exploratory Data Analysis                                              |     Shashwat      | Bharath  |
|                \-                 | Data cleaning and Data pre-processing based on EDA                     |      Bharath      | Pranshu  |
|                \-                 | Question specific exploration and identify initial trends and patterns |       Joel        |  Vishal  |
|     November 16^th^ to 22^nd^     | Model training for Q1                                                  |      Vishal       | Shashwat |
|                \-                 | Model training for Q2                                                  |      Pranshu      |   Joel   |
|     November 23^rd^ to 29^th^     | Continue Model training and testing for Q1 and Q2                      |      Vishal       | Pranshu  |
|                \-                 | Improving the models if there is a need                                |       Joel        | Bharath  |
| November 30^th^ to December 6^th^ | Refining the code for code review with comments                        |      Bharath      |  Vishal  |
|                \-                 | Generate insights from the model output                                |     Shashwat      |   Joel   |
|     December 7^th^ to 13^th^      | Review the generated models                                            |      Pranshu      | Shashwat |
|                \-                 | Write-up and presentation for the project                              |     Everyone      | Everyone |

### Repo Organization

The following are the folders involved in the Project repository.

-   **'data/':** Used for storing any necessary data files for the project, such as input files.

-   **'images/':** Used for storing image files used in the project.

-   **'presentation_files/':** Folder for having presentation related files.

-   **'\_extra/':** Used to brainstorm our analysis which won't impact our project workflow.

-   **'\_freeze/':** This folder is used to store the generated files during the build process. These files represent the frozen state of the website at a specific point in time.

-   **'\_site/':** Folder used to store the generated static website files after the site generator processes the quarto document.

-   **'.github/':** Folder for storing github templates and workflow.

We will be creating few folders inside `images/` folder for storing question specific images and presentation related images which are generated through out the plot. We will be creating `images/Q1`, `images/Q2` and `images/Presentation` for those respective files.

::: callout-note
## Note:

These are the planned approaches, and we intend to explore and solve the problem statement which we came up with. Parts of our approach might change in the final implementation.
:::
