---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Pattern Pioneers - "
    affiliations:
      - name: "School of Information, University of Arizona"
description: ""
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
categories: 
  - Data Mining
execute:
  warning: false
---

## Load required packages

```{r load_pkgs, message=FALSE, output=FALSE}
#| code-fold: true
#| code-summary: "Installed Packages"

### GETTING THE LIBRARIES
if (!require(pacman))
  install.packages(pacman)

pacman::p_load(tidyverse,
               tidytuesdayR,
               dlookr,
               formattable)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

## Examine data

Using dlookr's `describe()` and `diagnose()` - some basic EDA

## Questions

### Question 1

### Question 2

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

### Approach for question 1

### Approach for question 2

## Organization

### Plan of Attack

### Repo Organization
